---
title: "Dataset generation"
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output: html_document
---

# Import pkgs 

```{r}
if (!require(pacman)) install.packages("pacman")

pacman::p_load(
  broom, 
  tidyverse,
  glue,
  here,
  purrr,
  estimatr,
  effectsize,
  sandwich,
  ggrepel,
  patchwork,
  ggpubr, 
  corrplot,
  corrr,
  correlation, 
  tidycensus,
  callr,
  naniar,
  modelsummary,
  confintr, 
  tinytable, 
  vroom, 
  gt,
  DT,
  sf,
  tigris,
  statebins,
  RColorBrewer
)
```

```{r}
library("tidylog", warn.conflicts = FALSE)

devtools::install_github("graemeblair/stdidx")
library(stdidx)

options(es.use_symbols = TRUE) # get nice symbols

source(here("functions", "utils.r"))

ggplot2::theme_set(custom_theme())
```

# Import data 

```{r}
# MMA data 
## master file 
mbf <- vroom(here("raw_data", "irs_mbf.csv"))

## org activities from the websites and elsewhere 
org_activities <- vroom(here("raw_data", "irs_org_activities.csv"))

irs_activities <- vroom(here("raw_data","irs_nonweb_activities.csv"))

irs_activities <- irs_activities %>%
  replace_na(list(volunteer_text = 0, member_text = 0)) # without this, the following mutate won't work

## prediction labels
irs_pred <- vroom(here("raw_data", "predictions.csv")) %>%
  select(ein, predicted)
```

```{r}
## PO BOX status
po_status <- read_csv(here("raw_data", "irs_po_status.csv"))

## orgs geolocated to FIPS
irs_to_fips <- vroom(here("raw_data", "irs_to_fips.csv"))
```

"HCOST" = "Housing cost burden among households" 
"POV150" = Persons living below 150% of the poverty level
"CROWD" = "Crowding among housing units"
"AGE65" = "Persons aged 65 years or older"     
"SNGPNT" = "Single-parent households"
"BROAD" = "No broadband internet subscription among households"
"NOHSDP" = "No high school diploma among adults aged 25 years or older" 
"UNEMP" = "Unemployment among people 16 years and older in the labor force"
"REMNRTY" = "Persons of racial or ethnic minority status"

```{r}
# SDOH (2017-2021 ACS adapted by CDC)

## County 
sdoh_county <- read_csv(here("raw_data", "sdoh_county.csv"))

sdoh_county <- sdoh_county %>%
  select(StateAbbr, StateDesc, LocationID, MeasureID, Data_Value, TotalPopulation, Geolocation)

sdoh_county_wide <- sdoh_county %>%
  pivot_wider(names_from = MeasureID, 
              values_from = Data_Value)

## Zipcode
sdoh_zcta <- read_csv(here("raw_data", "sdoh_zcta.csv")) 

sdoh_zcta <- sdoh_zcta %>%
  select(LocationID, MeasureID, Data_Value, TotalPopulation, Geolocation)

sdoh_zcta_wide <- sdoh_zcta %>%
  pivot_wider(names_from = MeasureID, 
              values_from = Data_Value)

sdoh_county_wide %>%
  select(HCOST, POV150, CROWD, AGE65, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  datasummary_correlation(., method = fun)

sdoh_zcta_wide %>%
  select(HCOST, POV150, CROWD, AGE65, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  datasummary_correlation(., method = fun)
```

# Create a sampling frame

```{r}
irs_activities <- irs_activities %>% left_join(irs_pred)

org_activities_combined <- irs_activities %>%
  left_join(org_activities) %>%
  mutate(
    volunteer = ifelse(volunteer_text == 1, 1, volunteer),
    membership = ifelse(member_text == 1, 1, membership)
  )
```

```{r}
org_activities_mbf <- org_activities_combined %>%
  left_join(mbf %>% dplyr::select(ein, state, city, zip, irs_group)) %>%
  mutate(grouping_value = if_else(irs_group != "0000", irs_group, ein)) %>%
  left_join(po_status) %>%
  left_join(irs_to_fips)

nrow(org_activities_mbf) # 1,774,798, approximately 1.8M oranizatiions in the sampling frame
```

```{r}
org_activities_mbf <- org_activities_mbf %>%
  mutate(extractable = ifelse(!is.na(take_action) | !is.na(volunteer) | !is.na(membership) | !is.na(events) | !is.na(resources) | !is.na(advocacy) | !is.na(services) | !is.na(chapters) | !is.na(board) | !is.na(press) | !is.na(donations), 1, 0)) %>%
  dplyr::select(ein, state, city, zip, take_action, volunteer, membership, events, extractable, grouping_value, is_po, fips, predicted)
```

```{r}
org_activities_mbf <- org_activities_mbf %>%
  # filter(extractable == 1) %>% # this includes both websites and IRS
  filter(is_po != 1) # PO BOX

nrow(org_activities_mbf) # 1,363,701

org_activities_mbf <- org_activities_mbf %>%
  replace_na(list(take_action = 0, volunteer = 0, membership = 0, events = 0))

# the first 5 digits
org_activities_mbf$ZCTA <- substr(org_activities_mbf$zip, 1, 5)
```
  
# Datasets

- County 
- ZCTA

```{r}
org_activities_mbf <- org_activities_mbf %>%
  mutate(
    opp_binary = ifelse(volunteer == 1 | membership == 1 | take_action == 1 | events == 1, 1, 0),
    opp_mean = idx_mean(volunteer, membership, take_action, events)
    ) # average civic opportunity score at the organization level

civic_orgs <- org_activities_mbf %>%
  filter(opp_binary == 1)
```

## The first datasets (civic opportunity degree)

```{r}
# County 
cnts_counts <- org_activities_mbf %>%
  rename(FIPS = fips) %>%
  group_by(FIPS, grouping_value, state) %>%
  summarise(
            b = max(opp_binary, na.rm = T),
            m = max(opp_mean, na.rm = T),
            volunteer_m = max(volunteer, na.rm = T),
            membership_m = max(membership, na.rm = T),
            take_action_m = max(take_action, na.rm = T),
            events_m = max(events, na.rm = T)) %>%
  group_by(FIPS, state) %>%
  summarise(
    n = n(),
    civic_org_sum = sum(b),
    civic_opp_sum = sum(m),
    volunteer_sum = sum(volunteer_m), 
    membership_sum = sum(membership_m), 
    take_action_sum = sum(take_action_m), 
    events_sum = sum(events_m)
  )

cnts_counts_cov <- cnts_counts %>%
  inner_join(sdoh_county_wide %>%
              select(-StateDesc), 
            by = c("FIPS" = "LocationID",
                   "state" = "StateAbbr")) # n = 3,126, previously n = 3,281

cnts_counts_cov <- cnts_counts_cov %>%
  ungroup() %>%
  mutate(
         civic_org_sum_normalized = 100000 * civic_org_sum/TotalPopulation, 
         civic_opp_sum_normalized = 100000 * civic_opp_sum/TotalPopulation,
         civic_opp_index = dplyr::ntile(civic_opp_sum_normalized, 5),
         volunteer_sum_normalized = 100000 * volunteer_sum/TotalPopulation,
         membership_sum_normalized = 100000 * membership_sum/TotalPopulation,
         take_action_sum_normalized = 100000 * take_action_sum/TotalPopulation,
         events_sum_normalized = 100000 * events_sum/TotalPopulation
         )

round(mean(cnts_counts$civic_opp_sum == 0), 2) # 3% of the counties do not have any organization offering civic opportunity 
round(mean(cnts_counts_cov$civic_opp_sum == 0), 2) # 1% of the counties do not have any organization offering civic opportunity 
## ADD BASELINES 

# export
write_csv(cnts_counts_cov, here("data_outputs", 
                                "county_counts_cov.csv"))
```

```{r}
# ZCTA
zcta_counts <- org_activities_mbf %>%
  group_by(ZCTA, grouping_value, state) %>%
  summarise(m = max(opp_mean, na.rm = T),
            b = max(opp_binary, na.rm = T),
            volunteer_m = max(volunteer, na.rm = T),
            membership_m = max(membership, na.rm = T),
            take_action_m = max(take_action, na.rm = T),
            events_m = max(events, na.rm = T)) %>%
  group_by(ZCTA, state) %>%
  summarise(
    n = n(),
    civic_org_sum = sum(b),
    civic_opp_sum = sum(m),
    volunteer_sum = sum(volunteer_m), 
    membership_sum = sum(membership_m), 
    take_action_sum = sum(take_action_m), 
    events_sum = sum(events_m)
  )
```

```{r}
zcta_counts_cov <- zcta_counts %>%  
  inner_join(sdoh_zcta_wide, 
            by = c("ZCTA" = "LocationID")) # n = 28,008, previously n = 30,988
```

```{r}
zcta_counts_cov <- zcta_counts_cov %>%
  ungroup() %>%
  mutate(
         civic_org_sum_normalized = 100000 * civic_org_sum/TotalPopulation, 
         civic_opp_sum_normalized = 100000 * civic_opp_sum/TotalPopulation,
         civic_opp_index = dplyr::ntile(civic_opp_sum_normalized, 5),
         volunteer_sum_normalized = 100000 * volunteer_sum/TotalPopulation,
         membership_sum_normalized = 100000 * membership_sum/TotalPopulation,
         take_action_sum_normalized = 100000 * take_action_sum/TotalPopulation,
         events_sum_normalized = 100000 * events_sum/TotalPopulation
         )

round(mean(zcta_counts$civic_opp_sum == 0), 2) # 17% of the zipcodes do not have any organization offering civic opportunity 
round(mean(zcta_counts_cov$civic_opp_sum == 0), 2) # 15% of the zipcodes do not have any organization offering civic opportunity 

## ADD BASELINES 

# export
write_csv(zcta_counts_cov, here("data_outputs", 
                                "zcta_counts_cov.csv"))
```

## The second datasets (the civic opportunity kind)

```{r}
class_vec <- c("arts", "civic", "community", "econ", "education", "foundation", "health", "hobby", "housing", "professional", "religious", "research", "socialfraternal", "unions", "youth")

civic_orgs <- civic_orgs %>%
  filter(!is.na(predicted)) %>%
  mutate(class = case_match(predicted,
                            class_vec[1] ~ "Arts & Cultural",
                            class_vec[2] ~ "Political",
                            class_vec[3] ~ "Community",
                            class_vec[4] ~ "Economic",
                            class_vec[5] ~ "Education",
                            class_vec[6] ~ "Foundations",
                            class_vec[7] ~ "Healthcare",
                            class_vec[8] ~ "Hobby & Sports",
                            class_vec[9] ~ "Housing",
                            class_vec[10] ~ "Professional",
                            class_vec[11] ~ "Religious",
                            class_vec[12] ~ "Research & Think Tank",
                            class_vec[13] ~ "Social & Fraternal",
                            class_vec[14] ~ "Unions",
                            class_vec[15] ~ "Youth"))
  
  
# County

cnty_civic_org_type <- civic_orgs %>%
  group_by(fips, class) %>%
  count() %>%
  ungroup() %>%
  group_by(fips) %>%
  mutate(freq = n/sum(n)) %>%
  group_by(fips) %>%
  mutate(primary_org_cat = class[which.max(freq)]) %>%
  ungroup()

write_csv(cnty_civic_org_type, here("data_outputs", "cnty_civic_org_type.csv"))

# ZCTA

zcta_civic_org_type <- civic_orgs %>%
  group_by(ZCTA, class) %>%
  count() %>%
  ungroup() %>%
  group_by(ZCTA) %>%
  mutate(freq = n/sum(n)) %>%
  group_by(ZCTA) %>%
  mutate(primary_org_cat = class[which.max(freq)]) %>%
  ungroup()

write_csv(zcta_civic_org_type, here("data_outputs", "zcta_civic_org_type.csv"))
```