---
title: "Validation"
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output: html_document
---

# Import pkgs 

```{r}
if (!require(pacman)) install.packages("pacman")

pacman::p_load(
  boot,          # For bootstrapping
  paralle,       # For parallel processing
  scales,        # For labels
  arrow,          # For reading the Parquet file
  broom,        # Convert model outputs to tidy tibbles
  tidyverse,      # Core data science tools: dplyr, ggplot2, etc. 
  glue,           # String interpolation 
  here,           # For reproducible file paths 
  purrr,          # Functional programming 
  estimatr,       # Robust regression 
  # effectsize,   # Effect size computation 
  sandwich,       # Robust standard errors 
  # ggrepel,      # Improved label placement in ggplot2 
  patchwork,      # Combine multiple ggplots 
  # ggpubr,       # Publication-ready plots 
  # corrplot,     # Correlation matrix plots
  # corrr,        # Tidy correlation workflows 
  correlation,  # Bayesian/frequentist correlation tests 
  modelsummary,   # Summarizing model outputs
  confintr,       # Confidence intervals 
  tinytable,     # Lightweight table formatting 
  binom          # For Wilson 95% CIs in get_predicted_percent_ci()
)

library("tidylog", warn.conflicts = FALSE)

devtools::install_github("graemeblair/stdidx")
library(stdidx)

options(es.use_symbols = TRUE) # get nice symbols

source(here("functions", "utils.r"))

ggplot2::theme_set(custom_theme())
```

# Import data 

```{r}
# Inidivdual org level data 
org_df <- arrow::read_parquet(here("raw_data", "ind_org_df.parquet"))

# County 
cnts_counts_cov <- read_csv(here("data_outputs", "cnty_counts_cov.csv"))
cnts_civic_org_type <- read_csv(here("data_outputs", "cnty_civic_org_type.csv"))

# Zipcode
zcta_counts_cov <- read_csv(here("data_outputs", "zcta_counts_cov.csv"))
zcta_civic_org_type <- read_csv(here("data_outputs", "zcta_civic_org_type.csv"))

cnts_civic_org_type <- cnts_civic_org_type %>% left_join(cnts_counts_cov %>%
                                                           select(FIPS, POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY), by = c("FIPS" = "FIPS"))

zcta_civic_org_type <- zcta_civic_org_type %>% left_join(zcta_counts_cov %>%
                                                           select(ZCTA, POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY), by = c("ZCTA"))
```

# Preprocessing 

```{r}
# County 
cnts_counts_cov <- cnts_counts_cov %>%
  #select(POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))

# Zipcode
zcta_counts_cov <- zcta_counts_cov %>%
  #select(POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))
```

# Validation

## The org level data

```{r}
summary_df_long <- org_df %>%
  mutate(
    civic_score = ifelse(
      (is.na(volunteer) & is.na(membership) & is.na(take_action) & is.na(events)) |
      (rowSums(across(c(volunteer, membership, take_action, events)), na.rm = TRUE) == 0),
      NA,
      idx_mean(volunteer, membership, take_action, events)
    ),
    civic_score_group = case_when(
      is.na(civic_score)        ~ "No Civic Score", 
      civic_score < 0.5         ~ "Low Civic Score",
      civic_score >= 0.5        ~ "Median or High Civic Score"
    ),
    civic_score_group = factor(civic_score_group, levels = c(
      "No Civic Score", "Low Civic Score", "Median or High Civic Score"
    ))
  ) %>%
  group_by(civic_score_group) %>%
  summarize(
    asset   = list(bootstrap_median_ci(asset_amt)),
    income  = list(bootstrap_median_ci(income_amt)),
    revenue = list(bootstrap_median_ci(revenue_amt)),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(asset, income, revenue),
    names_to = "metric",
    values_to = "boot_result"
  ) %>%
  unnest_wider(boot_result, names_sep = "_")  # yields boot_result_median, _ci_low, _ci_high

unique(summary_df_long$civic_score_group) # should be three groups 
```

### Figure 4

```{r}
summary_df_long <- summary_df_long %>%
  mutate(metric = str_to_title(metric))  # Capitalize: "Asset", "Income", "Revenue")
```

```{r}
# Open PNG graphics device (base R)
png(filename = here::here("plots", "figure4.png"), width = 1200, height = 900, res = 120)

# Plot
ggplot(summary_df_long, aes(x = factor(civic_score_group), y = boot_result_median, group = metric)) +
  geom_col(aes(fill = metric), position = position_dodge(width = 0.8), width = 0.7, alpha = 0.7) +
  geom_errorbar(
    aes(ymin = boot_result_ci_low, ymax = boot_result_ci_high, color = metric),
    width = 0.3,
    linewidth = 1.5,  # Make thicker
    position = position_dodge(width = 0.8)
  ) +
  geom_point(
    aes(color = metric),
    position = position_dodge(width = 0.8),
    size = 2
  ) +
  geom_text(
    aes(label = scales::label_dollar()(boot_result_median)),  # No scale or suffix
    position = position_dodge(width = 0.8),
    vjust = -0.5,
    size = 4
  ) +
  labs(
    x = "Civic opportunity score",
    y = "Median amount (excluding NAs and 0s)",
    fill = "Financial Metric",
    color = "Financial Metric",
    title = "Median Asset, Income, and Revenue by Civic Opportunity Score",
    subtitle = "95% confidence intervals are created by bootstrapping"
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(),  # Full dollar format, e.g., $125,000
    expand = expansion(mult = c(0, 0.1))
  ) +
  theme_minimal(base_size = 14) +
  guides(color = guide_legend(override.aes = list(fill = NA)))  # Combine color & fill legends

# Close PNG device
dev.off()
```

## The first dataset (degree)

"POV150" = "Persons living below 150% of the poverty level"
"SNGPNT" = "Single-parent households"
"BROAD" = "No broadband internet subscription among households"
"NOHSDP" = "No high school diploma among adults aged 25 years or older" 
"UNEMP" = "Unemployment among people 16 years and older in the labor force"
"REMNRTY" = "Persons of racial or ethnic minority status"

### County 

#### Figure 5

```{r}
wrap_plots(
  cnty_pred_plot(REMNRTY, "Persons of racial or ethnic minority status"), # race
  cnty_pred_plot(POV150, "Persons living below 150% of the poverty level"), # poverty
  cnty_pred_plot(NOHSDP, "No high school diploma among adults aged 25 years or older"), # college education
  cnty_pred_plot(SNGPNT, "Single-parent households"), # single parent household 
  cnty_pred_plot(UNEMP, "Unemployment among people 16 years and older"), # unemployment 
  cnty_pred_plot(BROAD, "No broadband internet subscription among households"), # internet connection
  ncol = 3
) +
  plot_annotation(tag_levels = "a", title = "Unit: County") +
  plot_layout(guides = "collect")

ggsave(here("plots", "figure5.png"), width = 14, height = 12)
```

### ZIP code 

#### Figure 6

```{r}
wrap_plots(
  zcta_pred_plot(REMNRTY, "Persons of racial or ethnic minority status"), # race
  zcta_pred_plot(POV150, "Persons living below 150% of the poverty level"), # poverty
  zcta_pred_plot(NOHSDP, "No high school diploma among adults aged 25 years or older"), # college education
  zcta_pred_plot(SNGPNT, "Single-parent households"), # single parent household 
  zcta_pred_plot(UNEMP, "Unemployment among people 16 years and older"), # unemployment 
  zcta_pred_plot(BROAD, "No broadband internet subscription among households"), # internet connection
  ncol = 3
) +
  plot_annotation(tag_levels = "a", title = "Unit: ZIP code") +
  plot_layout(guides = "collect")

ggsave(here("plots", "figure6.png"), width = 14, height = 12)
```

## The second dataset (kind)

### County 

```{r}
cnts_civic_org_type <- cnts_civic_org_type %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))

cnts_civic_org_type <- cnts_civic_org_type %>%
  mutate(new_class = ifelse(str_detect(class, "Rel|Social"), "Religious or Social & Fraternal", "Others"))

cnts_civic_org_new_type <- cnts_civic_org_type %>%
  group_by(new_class, FIPS) %>%
  mutate(re_or_so_freq = sum(freq)) %>%
  ungroup() %>%
  select(FIPS, re_or_so_freq, POV150:REMNRTY) %>%
  distinct()

sum_cnts_civic_org_type <- summary_class_civic_org(cnts_civic_org_new_type)
  
cnty_type_plot <- sum_cnts_civic_org_type %>%
  ggplot(aes(x = fct_reorder(name, estimate), y = estimate, ymax = conf.high, ymin = conf.low)) +
  geom_col(alpha = 0.5, position = position_dodge(width = 0.9)) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.3) +
  coord_flip() +
  labs(
    title = "Unit: County",
    x = "",
    y = "Estimated OLS Coefficients",
    fill = "Civic opportunity organization type") +
  geom_hline(yintercept = 0, col = "red", linetype = 2) 
```

### ZIP code 

```{r}
zcta_civic_org_type <- zcta_civic_org_type %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))

zcta_civic_org_type <- zcta_civic_org_type %>%
  mutate(new_class = ifelse(str_detect(class, "Rel|Social"), "Religious or Social & Fraternal", "Others"))

zcta_civic_org_new_type <- zcta_civic_org_type %>%
  group_by(new_class, ZCTA) %>%
  mutate(re_or_so_freq = sum(freq)) %>%
  ungroup() %>%
  select(ZCTA, re_or_so_freq, POV150:REMNRTY) %>%
  distinct()

sum_zcta_civic_org_type <- summary_class_civic_org(zcta_civic_org_new_type)
```

```{r}
zcta_type_plot <- sum_zcta_civic_org_type %>%
 ggplot(aes(x = fct_reorder(name, estimate), y = estimate, ymax = conf.high, ymin = conf.low)) +
  geom_col(alpha = 0.5, position = position_dodge(width = 0.9)) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.3) +
  coord_flip() +
  labs(
    title = "Unit: ZIP code",
    x = "",
    y = "Estimated OLS Coefficients",
    fill = "Civic opportunity organization type") +
  geom_hline(yintercept = 0, col = "red", linetype = 2)
```

```{r}
binded_sums <- bind_rows(sum_cnts_civic_org_type %>% 
            mutate(unit = "County"), 
                         sum_zcta_civic_org_type %>%
            mutate(unit = "ZIP code"))
```

### Figure 7

```{r}
binded_sums %>%
  ggplot(aes(x = fct_reorder(name, estimate), y = estimate, ymax = conf.high, ymin = conf.low)) +
  geom_col(alpha = 0.5, position = position_dodge(width = 0.9)) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.3) +
  coord_flip() +
  labs(
    title = "Civic opportunity dependendcy on religious or social fraternal organizations",
    x = "",
    y = "Estimated OLS Coefficients",
    fill = "Civic opportunity organization type") +
  geom_hline(yintercept = 0, col = "red", linetype = 2) +
  facet_wrap(~unit)

ggsave(here("plots", "figure7.png"), width = 10)
```