---
title: "Validation"
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output: html_document
---

# Import pkgs 

```{r}
if (!require(pacman)) install.packages("pacman")

pacman::p_load(
  broom, 
  tidyverse,
  glue,
  here,
  purrr,
  estimatr,
  effectsize,
  sandwich,
  ggrepel,
  patchwork,
  ggpubr, 
  corrplot,
  corrr,
  correlation, 
  modelsummary,
  confintr, 
  tinytable
)

library("tidylog", warn.conflicts = FALSE)

devtools::install_github("graemeblair/stdidx")
library(stdidx)

options(es.use_symbols = TRUE) # get nice symbols

source(here("functions", "utils.r"))

ggplot2::theme_set(custom_theme())
```

# Import data 

```{r}
# County 
cnts_counts_cov <- read_csv(here("data_outputs", "county_counts_cov.csv"))
cnts_civic_org_type <- read_csv(here("data_outputs", "cnty_civic_org_type.csv"))

# Zipcode
zcta_counts_cov <- read_csv(here("data_outputs", "zcta_counts_cov.csv"))
zcta_civic_org_type <- read_csv(here("data_outputs", "zcta_civic_org_type.csv"))

cnts_civic_org_type <- cnts_civic_org_type %>% left_join(cnts_counts_cov %>%
                                                           select(FIPS, POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY), by = c("FIPS" = "FIPS"))

zcta_civic_org_type <- zcta_civic_org_type %>% left_join(zcta_counts_cov %>%
                                                           select(ZCTA, POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY), by = c("ZCTA"))
```

# Preprocessing 

```{r}
# County 
cnts_counts_cov <- cnts_counts_cov %>%
  #select(POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))

# Zipcode
zcta_counts_cov <- zcta_counts_cov %>%
  #select(POV150, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))
```

# Validation

## The first dataset (degree)

"POV150" = "Persons living below 150% of the poverty level"
"SNGPNT" = "Single-parent households"
"BROAD" = "No broadband internet subscription among households"
"NOHSDP" = "No high school diploma among adults aged 25 years or older" 
"UNEMP" = "Unemployment among people 16 years and older in the labor force"
"REMNRTY" = "Persons of racial or ethnic minority status"

### County 

#### Figure 4

```{r}
wrap_plots(
  cnty_pred_plot(REMNRTY, "Persons of racial or ethnic minority status"), # race
  cnty_pred_plot(POV150, "Persons living below 150% of the poverty level"), # poverty
  cnty_pred_plot(NOHSDP, "No high school diploma among adults aged 25 years or older"), # college education
  cnty_pred_plot(SNGPNT, "Single-parent households"), # single parent household 
  cnty_pred_plot(UNEMP, "Unemployment among people 16 years and older"), # unemployment 
  cnty_pred_plot(BROAD, "No broadband internet subscription among households"), # internet connection
  ncol = 3
) +
  plot_annotation(tag_levels = "a", title = "Unit: County") +
  plot_layout(guides = "collect")

ggsave(here("plots", "figure4.png"), width = 14, height = 12)
```

### ZIP code 

#### Figure 5

```{r}
wrap_plots(
  zcta_pred_plot(REMNRTY, "Persons of racial or ethnic minority status"), # race
  zcta_pred_plot(POV150, "Persons living below 150% of the poverty level"), # poverty
  zcta_pred_plot(NOHSDP, "No high school diploma among adults aged 25 years or older"), # college education
  zcta_pred_plot(SNGPNT, "Single-parent households"), # single parent household 
  zcta_pred_plot(UNEMP, "Unemployment among people 16 years and older"), # unemployment 
  zcta_pred_plot(BROAD, "No broadband internet subscription among households"), # internet connection
  ncol = 3
) +
  plot_annotation(tag_levels = "a", title = "Unit: ZIP code") +
  plot_layout(guides = "collect")

ggsave(here("plots", "figure5.png"), width = 14, height = 12)
```

## The second dataset (kind)

### County 

```{r}
cnts_civic_org_type <- cnts_civic_org_type %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))

cnts_civic_org_type <- cnts_civic_org_type %>%
  mutate(new_class = ifelse(str_detect(class, "Rel|Social"), "Religious or Social & Fraternal", "Others"))

cnts_civic_org_new_type <- cnts_civic_org_type %>%
  group_by(new_class, FIPS) %>%
  mutate(re_or_so_freq = sum(freq)) %>%
  ungroup() %>%
  select(FIPS, re_or_so_freq, POV150:REMNRTY) %>%
  distinct()

sum_cnts_civic_org_type <- summary_class_civic_org(cnts_civic_org_new_type)
  
cnty_type_plot <- sum_cnts_civic_org_type %>%
  ggplot(aes(x = fct_reorder(name, estimate), y = estimate, ymax = conf.high, ymin = conf.low)) +
  geom_col(alpha = 0.5, position = position_dodge(width = 0.9)) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.3) +
  coord_flip() +
  labs(
    title = "Unit: County",
    x = "",
    y = "Estimated OLS Coefficients",
    fill = "Civic opportunity organization type") +
  geom_hline(yintercept = 0, col = "red", linetype = 2) 
```

### ZIP code 

```{r}
zcta_civic_org_type <- zcta_civic_org_type %>%
  mutate(POV150 = normalize(POV150),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))

zcta_civic_org_type <- zcta_civic_org_type %>%
  mutate(new_class = ifelse(str_detect(class, "Rel|Social"), "Religious or Social & Fraternal", "Others"))

zcta_civic_org_new_type <- zcta_civic_org_type %>%
  group_by(new_class, ZCTA) %>%
  mutate(re_or_so_freq = sum(freq)) %>%
  ungroup() %>%
  select(ZCTA, re_or_so_freq, POV150:REMNRTY) %>%
  distinct()

sum_zcta_civic_org_type <- summary_class_civic_org(zcta_civic_org_new_type)
```

```{r}
zcta_type_plot <- sum_zcta_civic_org_type %>%
 ggplot(aes(x = fct_reorder(name, estimate), y = estimate, ymax = conf.high, ymin = conf.low)) +
  geom_col(alpha = 0.5, position = position_dodge(width = 0.9)) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.3) +
  coord_flip() +
  labs(
    title = "Unit: ZIP code",
    x = "",
    y = "Estimated OLS Coefficients",
    fill = "Civic opportunity organization type") +
  geom_hline(yintercept = 0, col = "red", linetype = 2)
```

```{r}
binded_sums <- bind_rows(sum_cnts_civic_org_type %>% 
            mutate(unit = "County"), 
                         sum_zcta_civic_org_type %>%
            mutate(unit = "ZIP code"))
```

### Figure 6 

```{r}
binded_sums %>%
  ggplot(aes(x = fct_reorder(name, estimate), y = estimate, ymax = conf.high, ymin = conf.low)) +
  geom_col(alpha = 0.5, position = position_dodge(width = 0.9)) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.3) +
  coord_flip() +
  labs(
    title = "Civic opportunity dependendcy on religious or social fraternal organizations",
    x = "",
    y = "Estimated OLS Coefficients",
    fill = "Civic opportunity organization type") +
  geom_hline(yintercept = 0, col = "red", linetype = 2) +
  facet_wrap(~unit)

ggsave(here("plots", "figure6.png"), width = 10)
```