---
title: "Validation"
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output: html_document
---

# Import pkgs 

```{r}
if (!require(pacman)) install.packages("pacman")

pacman::p_load(
  broom, 
  tidyverse,
  glue,
  here,
  purrr,
  estimatr,
  effectsize,
  sandwich,
  ggrepel,
  patchwork,
  ggpubr, 
  corrplot,
  corrr,
  correlation, 
  tidycensus,
  callr,
  naniar,
  modelsummary,
  confintr, 
  tinytable, 
  vroom, 
  gt,
  DT,
  sf,
  tigris,
  statebins,
  RColorBrewer
)

library("tidylog", warn.conflicts = FALSE)

devtools::install_github("graemeblair/stdidx")
library(stdidx)

options(es.use_symbols = TRUE) # get nice symbols

source(here("functions", "utils.r"))

ggplot2::theme_set(custom_theme())
```

# Import data 

```{r}
# County 
cnts_counts_cov <- read_csv(here("data_outputs", "county_counts_cov.csv"))
cnts_civic_org_type <- read_csv(here("data_outputs", "cnty_civic_org_type.csv"))

# Zipcode
zcta_counts_cov <- read_csv(here("data_outputs", "zcta_counts_cov.csv"))
zcta_civic_org_type <- read_csv(here("data_outputs", "zcta_civic_org_type.csv"))

cnts_civic_org_type <- cnts_civic_org_type %>% left_join(cnts_counts_cov %>%
                                                           select(FIPS, HCOST, POV150, CROWD, AGE65, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY), by = c("fips" = "FIPS"))

zcta_civic_org_type <- zcta_civic_org_type %>% left_join(zcta_counts_cov %>%
                                                           select(ZCTA, HCOST, POV150, CROWD, AGE65, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY), by = c("ZCTA"))
```

# Preprocessing 

```{r}
# County 
cnts_counts_cov <- cnts_counts_cov %>%
  #select(HCOST, POV150, CROWD, AGE65, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  mutate(HCOST = normalize(HCOST), 
         POV150 = normalize(POV150),
         CROWD = normalize(CROWD),
         AGE65 = normalize(AGE65),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))

# Zipcode
zcta_counts_cov <- zcta_counts_cov %>%
  #select(HCOST, POV150, CROWD, AGE65, SNGPNT, BROAD, NOHSDP, UNEMP, REMNRTY) %>%
  mutate(HCOST = normalize(HCOST), 
         POV150 = normalize(POV150),
         CROWD = normalize(CROWD),
         AGE65 = normalize(AGE65),
         SNGPNT = normalize(SNGPNT),
         BROAD = normalize(BROAD),
         NOHSDP = normalize(NOHSDP),
         UNEMP = normalize(UNEMP),
         REMNRTY = normalize(REMNRTY))
```

# Validation

## The first dataset (degree)

"HCOST" = "Housing cost burden among households" 
"POV150" = "Persons living below 150% of the poverty level"
"CROWD" = "Crowding among housing units"
"AGE65" = "Persons aged 65 years or older"     
"SNGPNT" = "Single-parent households"
"BROAD" = "No broadband internet subscription among households"
"NOHSDP" = "No high school diploma among adults aged 25 years or older" 
"UNEMP" = "Unemployment among people 16 years and older in the labor force"
"REMNRTY" = "Persons of racial or ethnic minority status"

```{r}
# County 

wrap_plots(
  cnty_pred_plot(POV150, "Persons living below 150% of the poverty level"), # poverty
  cnty_pred_plot(NOHSDP, "No high school diploma among adults aged 25 years or older"), # college education
  cnty_pred_plot(REMNRTY, "Persons of racial or ethnic minority status"), # race
  cnty_pred_plot(AGE65, "Persons aged 65 years or older"), # age
  cnty_pred_plot(SNGPNT, "Single-parent households"), # household conditions 
  cnty_pred_plot(UNEMP, "Unemployment among people 16 years and older"), # household conditions 
   # infrastructure 
  cnty_pred_plot(HCOST, "Housing cost burden among households"), # housing conditions 
  cnty_pred_plot(CROWD, "Crowding among housing units"), # housing conditions 
  cnty_pred_plot(BROAD, "No broadband internet subscription among households"), # neighborhood conditions 
  ncol = 3
) +
  plot_annotation(tag_levels = "a", title = "Unit: County") +
  plot_layout(guides = "collect")

ggsave(here("plots", "construct_validity_county.png"), width = 14, height = 14)
```


```{r}
# Zipcode

wrap_plots(
  zcta_pred_plot(POV150, "Persons living below 150% of the poverty level"), # poverty
  zcta_pred_plot(NOHSDP, "No high school diploma among adults aged 25 years or older"), # college education
  zcta_pred_plot(REMNRTY, "Persons of racial or ethnic minority status"), # race
  zcta_pred_plot(AGE65, "Persons aged 65 years or older"), # age
  zcta_pred_plot(SNGPNT, "Single-parent households"), # household conditions 
  zcta_pred_plot(UNEMP, "Unemployment among people 16 years and older"), # household conditions 
   # infrastructure 
  zcta_pred_plot(HCOST, "Housing cost burden among households"), # housing conditions 
  zcta_pred_plot(CROWD, "Crowding among housing units"), # housing conditions 
  zcta_pred_plot(BROAD, "No broadband internet subscription among households"), # neighborhood conditions 
  ncol = 3
) +
  plot_annotation(tag_levels = "a", title = "Unit: ZIP code") +
  plot_layout(guides = "collect")

ggsave(here("plots", "construct_validity_zcta.png"), width = 14, height = 14)
```

## The second dataset 

```{r}
# County
cnts_civic_org_type <- cnts_civic_org_type %>%
   mutate(HCOST = normalize(HCOST), 
          POV150 = normalize(POV150),
          CROWD = normalize(CROWD),
          AGE65 = normalize(AGE65),
          SNGPNT = normalize(SNGPNT),
          BROAD = normalize(BROAD),
          NOHSDP = normalize(NOHSDP),
          UNEMP = normalize(UNEMP),
          REMNRTY = normalize(REMNRTY))

sum_cnts_civic_org_type <- reduce(list(summary_primary_civic_org(cnts_civic_org_type, POV150, "Living below 150% of the poverty level"),
summary_primary_civic_org(cnts_civic_org_type, NOHSDP, "No high school diploma among 25 yrs+"),
summary_primary_civic_org(cnts_civic_org_type, REMNRTY, "Persons of racial or ethnic minority status"),
summary_primary_civic_org(cnts_civic_org_type, AGE65, "Persons aged 65 yrs+"),
summary_primary_civic_org(cnts_civic_org_type, SNGPNT, "Single-parent households"),
summary_primary_civic_org(cnts_civic_org_type, UNEMP, "Unemployment among people 16 yrs+"),
summary_primary_civic_org(cnts_civic_org_type, HCOST, "Housing cost burden among households"),
summary_primary_civic_org(cnts_civic_org_type, CROWD, "Crowding among housing units"),
summary_primary_civic_org(cnts_civic_org_type, BROAD, "No broadband internet subscription")), bind_rows)

sum_cnts_civic_org_type %>%
  ggplot(aes(x = fct_reorder(primary_org_cat, avg), y = avg, ymax = hi_ci, ymin = low_ci)) +
  geom_col(alpha = 0.2) +
  geom_errorbar() +
  facet_wrap(~variable, ncol = 3) +
  coord_flip() +
  labs(x = "Primary civic opportuntiy provider") +
  scale_y_continuous(label = scales::percent) +
  geom_hline(yintercept = 0.5, col = "red", linetype = 2)

ggsave(here("plots", "civic_opp_provider_cnty.png"), width = 12, height = 12)
```

```{r}
# ZCTA
zcta_civic_org_type <- zcta_civic_org_type %>%
   mutate(HCOST = normalize(HCOST), 
          POV150 = normalize(POV150),
          CROWD = normalize(CROWD),
          AGE65 = normalize(AGE65),
          SNGPNT = normalize(SNGPNT),
          BROAD = normalize(BROAD),
          NOHSDP = normalize(NOHSDP),
          UNEMP = normalize(UNEMP),
          REMNRTY = normalize(REMNRTY))

sum_zcta_civic_org_type <- reduce(list(summary_primary_civic_org(zcta_civic_org_type, POV150, "Living below 150% of the poverty level"),
summary_primary_civic_org(zcta_civic_org_type, NOHSDP, "No high school diploma among 25 yrs+"),
summary_primary_civic_org(zcta_civic_org_type, REMNRTY, "Persons of racial or ethnic minority status"),
summary_primary_civic_org(zcta_civic_org_type, AGE65, "Persons aged 65 yrs+"),
summary_primary_civic_org(zcta_civic_org_type, SNGPNT, "Single-parent households"),
summary_primary_civic_org(zcta_civic_org_type, UNEMP, "Unemployment among people 16 yrs+"),
summary_primary_civic_org(zcta_civic_org_type, HCOST, "Housing cost burden among households"),
summary_primary_civic_org(zcta_civic_org_type, CROWD, "Crowding among housing units"),
summary_primary_civic_org(zcta_civic_org_type, BROAD, "No broadband internet subscription")), bind_rows)

sum_zcta_civic_org_type %>%
  ggplot(aes(x = fct_reorder(primary_org_cat, avg), y = avg, ymax = hi_ci, ymin = low_ci)) +
  geom_col(alpha = 0.2) +
  geom_errorbar() +
  facet_wrap(~variable, ncol = 3) +
  coord_flip() +
  labs(x = "Primary civic opportuntiy provider") +
  scale_y_continuous(label = scales::percent) +
  geom_hline(yintercept = 0.5, col = "red", linetype = 2)

ggsave(here("plots", "civic_opp_provider_zcta.png"), width = 12, height = 12)
```