---
title: "Data wrangling"
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output: html_document
---

# Import pkgs 

```{r}
if (!require(pacman)) install.packages("pacman")

pacman::p_load(
  broom, 
  tidyverse,
  glue,
  here,
  purrr,
  estimatr,
  effectsize,
  sandwich,
  ggrepel,
  patchwork,
  ggpubr, 
  corrplot,
  corrr,
  correlation, 
  tidycensus,
  callr,
  naniar,
  modelsummary,
  confintr, 
  tinytable, 
  vroom, 
  gt,
  DT,
  sf,
  tigris,
  statebins,
  RColorBrewer
)
```

```{r}

library("tidylog", warn.conflicts = FALSE)

devtools::install_github("graemeblair/stdidx")
library(stdidx)

options(es.use_symbols = TRUE) # get nice symbols

custom_theme <- function(size = 13) {
  theme_bw(base_size = size) +
    theme(
      aspect.ratio = 1.2,
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      axis.text.x = element_text(margin = margin(t = 6)),
      plot.title = element_text(size = 12),
      plot.subtitle = element_text(size = 10),
      plot.caption = element_text(colour = "grey50", hjust = 0),
      legend.position = "bottom"
    )
}

ggplot2::theme_set(custom_theme())
```

# Import data 

```{r}
# MMA data 
## master file 
mbf <- vroom(here("raw_data", "irs_mbf.csv"))

## org activities from the websites and elsewhere 
org_activities <- vroom(here("raw_data", "irs_org_activities.csv"))

irs_activities <- vroom(here("raw_data","irs_nonweb_activities.csv"))

irs_activities <- irs_activities %>%
  replace_na(list(volunteer_text = 0, member_text = 0)) # without this, the following mutate won't work

## prediction labels
irs_pred <- vroom(here("raw_data", "predictions.csv")) %>%
  select(ein, predicted)
```

```{r}
## PO BOX status
po_status <- read_csv(here("raw_data", "irs_po_status.csv"))

## orgs geolocated to FIPS
irs_to_fips <- vroom(here("raw_data", "irs_to_fips.csv"))
```

"HCOST" = "Housing cost burden among households" 
"POV150" = Persons living below 150% of the poverty level
"CROWD" = "Crowding among housing units"
"AGE65" = "Persons aged 65 years or older"     
"SNGPNT" = "Single-parent households"
"BROAD" = "No broadband internet subscription among households"
"NOHSDP" = "No high school diploma among adults aged 25 years or older" 
"UNEMP" = "Unemployment among people 16 years and older in the labor force"
"REMNRTY" = "Persons of racial or ethnic minority status"

```{r}
# SDOH (2017-2021 ACS adapted by CDC)

## County 
sdoh_county <- read_csv(here("raw_data", "sdoh_county.csv"))

sdoh_county <- sdoh_county %>%
  select(StateAbbr, StateDesc, LocationID, MeasureID, Data_Value, TotalPopulation, Geolocation)

sdoh_county_wide <- sdoh_county %>%
  pivot_wider(names_from = MeasureID, 
              values_from = Data_Value)

## Zipcode
sdoh_zcta <- read_csv(here("raw_data", "sdoh_zcta.csv")) 

sdoh_zcta <- sdoh_zcta %>%
  select(LocationName, MeasureID, Data_Value, TotalPopulation, Geolocation)

sdoh_zcta_wide <- sdoh_zcta %>%
  pivot_wider(names_from = MeasureID, 
              values_from = Data_Value)
```

# Create a sampling frame

```{r}
irs_activities <- irs_activities %>% left_join(irs_pred)

org_activities_combined <- irs_activities %>%
  left_join(org_activities) %>%
  mutate(
    volunteer = ifelse(volunteer_text == 1, 1, volunteer),
    membership = ifelse(member_text == 1, 1, membership)
  )
```

```{r}
org_activities_mbf <- org_activities_combined %>%
  left_join(mbf %>% dplyr::select(ein, state, city, zip, irs_group)) %>%
  mutate(grouping_value = if_else(irs_group != "0000", irs_group, ein)) %>%
  left_join(po_status) %>%
  left_join(irs_to_fips)

nrow(org_activities_mbf) # 1,774,798, approximately 1.8M oranizatiions in the sampling frame
```

```{r}
org_activities_mbf <- org_activities_mbf %>%
  mutate(extractable = ifelse(!is.na(take_action) | !is.na(volunteer) | !is.na(membership) | !is.na(events) | !is.na(resources) | !is.na(advocacy) | !is.na(services) | !is.na(chapters) | !is.na(board) | !is.na(press) | !is.na(donations), 1, 0)) %>%
  dplyr::select(ein, state, city, zip, take_action, volunteer, membership, events, extractable, grouping_value, is_po, fips, predicted)
```

```{r}
org_activities_mbf <- org_activities_mbf %>%
  # filter(extractable == 1) %>% # this includes both websites and IRS
  filter(is_po != 1) # PO BOX

nrow(org_activities_mbf) # 1,363,701

org_activities_mbf <- org_activities_mbf %>%
  replace_na(list(take_action = 0, volunteer = 0, membership = 0, events = 0))

# the first 5 digits
org_activities_mbf$ZCTA <- substr(org_activities_mbf$zip, 1, 5)
```

```{r}
nonweb_eins <- irs_activities %>%
  rename(
    volunteer = volunteer_text,
    membership = member_text
  ) %>%
  select(ein, volunteer, membership) %>%
  # replace_na(list(volunteer = 0, membership = 0)) %>%
  filter(volunteer == 1 | membership == 1) %>%
  pull(ein)

org_activities_mbf$nonweb <- org_activities_mbf$ein %in% nonweb_eins

org_activities_mbf %>%
  mutate(web = ifelse(nonweb == FALSE | nonweb == TRUE & take_action == 1 | nonweb == TRUE & events == 1, TRUE, FALSE)) %>%
  mutate(status = as.numeric(nonweb) + as.numeric(web)) %>%
  mutate(status = ifelse(status == 2, "Both",
    ifelse(status == 1 & web == TRUE, "Web",
      ifelse(status == 1 & nonweb == TRUE, "IRS", NA)
    )
  )) %>%
  group_by(status) %>%
  summarize(n = n()) %>%
  mutate(freq = n / sum(n))
```
  
# Datasets

- County 
- ZCTA

```{r}
org_activities_mbf <- org_activities_mbf %>%
  mutate(
    opp_binary = ifelse(volunteer == 1 | membership == 1 | take_action == 1 | events == 1, 1, 0),
    opp_mean = idx_mean(volunteer, membership, take_action, events)
    ) # average civic opportunity score at the organization level
```

```{r}
# County 
cnts_counts <- org_activities_mbf %>%
  rename(FIPS = fips) %>%
  group_by(FIPS, grouping_value, state) %>%
  summarise(
            b = max(opp_binary, na.rm = T),
            m = max(opp_mean, na.rm = T),
            volunteer_m = max(volunteer, na.rm = T),
            membership_m = max(membership, na.rm = T),
            take_action_m = max(take_action, na.rm = T),
            events_m = max(events, na.rm = T)) %>%
  group_by(FIPS, state) %>%
  summarise(
    n = n(),
    civic_org_sum = sum(b),
    civic_opp_sum = sum(m),
    volunteer_sum = sum(volunteer_m), 
    membership_sum = sum(membership_m), 
    take_action_sum = sum(take_action_m), 
    events_sum = sum(events_m)
  )

cnts_counts_cov <- cnts_counts %>%
  inner_join(sdoh_county_wide %>%
              select(-StateDesc), 
            by = c("FIPS" = "LocationID",
                   "state" = "StateAbbr")) # n = 3,126, previously n = 3,281

cnts_counts_cov <- cnts_counts_cov %>%
  ungroup() %>%
  mutate(
         civic_org_sum_normalized = 100000 * civic_org_sum/TotalPopulation, 
         civic_opp_sum_normalized = 100000 * civic_opp_sum/TotalPopulation,
         civic_opp_index = dplyr::ntile(civic_opp_sum_normalized, 5),
         volunteer_sum_normalized = 100000 * volunteer_sum/TotalPopulation,
         membership_sum_normalized = 100000 * membership_sum/TotalPopulation,
         take_action_sum_normalized = 100000 * take_action_sum/TotalPopulation,
         events_sum_normalized = 100000 * events_sum/TotalPopulation
         )

round(mean(cnts_counts$civic_opp_sum == 0), 2) # 3% of the counties do not have any organization offering civic opportunity 
round(mean(cnts_counts_cov$civic_opp_sum == 0), 2) # 1% of the counties do not have any organization offering civic opportunity 

# export
write_csv(cnts_counts_cov, here("data_outputs", 
                                "county_counts_cov.csv"))
```

```{r}
# ZCTA
zcta_counts <- org_activities_mbf %>%
  group_by(ZCTA, grouping_value, state) %>%
  summarise(m = max(opp_mean, na.rm = T),
            b = max(opp_binary, na.rm = T),
            volunteer_m = max(volunteer, na.rm = T),
            membership_m = max(membership, na.rm = T),
            take_action_m = max(take_action, na.rm = T),
            events_m = max(events, na.rm = T)) %>%
  group_by(ZCTA, state) %>%
  summarise(
    n = n(),
    civic_org_sum = sum(b),
    civic_opp_sum = sum(m),
    volunteer_sum = sum(volunteer_m), 
    membership_sum = sum(membership_m), 
    take_action_sum = sum(take_action_m), 
    events_sum = sum(events_m)
  )

zcta_counts_cov <- zcta_counts %>%  
  inner_join(sdoh_zcta_wide, 
            by = c("ZCTA" = "LocationName")) # n = 28,008, previously n = 30,988

zcta_counts_cov <- zcta_counts_cov %>%
  ungroup() %>%
  mutate(
         civic_org_sum_normalized = 100000 * civic_org_sum/TotalPopulation, 
         civic_opp_sum_normalized = 100000 * civic_opp_sum/TotalPopulation,
         civic_opp_index = dplyr::ntile(civic_opp_sum_normalized, 5),
         volunteer_sum_normalized = 100000 * volunteer_sum/TotalPopulation,
         membership_sum_normalized = 100000 * membership_sum/TotalPopulation,
         take_action_sum_normalized = 100000 * take_action_sum/TotalPopulation,
         events_sum_normalized = 100000 * events_sum/TotalPopulation
         )

round(mean(zcta_counts$civic_opp_sum == 0), 2) # 17% of the zipcodes do not have any organization offering civic opportunity 
round(mean(zcta_counts_cov$civic_opp_sum == 0), 2) # 15% of the zipcodes do not have any organization offering civic opportunity 

# export
write_csv(zcta_counts_cov, here("data_outputs", 
                                "zcta_counts_cov.csv"))
```

# Description

```{r}
# County 
datasummary(civic_org_sum_normalized + volunteer_sum_normalized + membership_sum_normalized + take_action_sum_normalized + events_sum_normalized ~ mean_no_na + std_no_na, 
            data = cnts_counts_cov, 
            output = "tinytable",
            fmt = 2) %>%
  save_tt(
    here("tables", "county_civic_opp_table.tex"),
    overwrite = TRUE
    )
```

```{r}
cnts_state_avg_se <- cnts_counts_cov %>%
  group_by(state) %>%
  summarize(avg = mean_no_na(civic_opp_index),
            se = std_no_na(civic_opp_index)) %>%
  mutate(unit = "County")
```

```{r}
# ZCTA
datasummary(civic_org_sum_normalized + volunteer_sum_normalized + membership_sum_normalized + take_action_sum_normalized + events_sum_normalized ~ mean_no_na + std_no_na, 
            data = zcta_counts_cov,
            output = "tinytable",
            fmt = 2) %>%
  save_tt(
    here("tables", "zcta_civic_opp_table.tex"),
    overwrite = TRUE
    )
```

```{r}
zcta_state_avg_se <- zcta_counts_cov %>%
  group_by(state) %>%
  summarize(avg = mean_no_na(civic_opp_index),
            se = std_no_na(civic_opp_index)) %>%
  mutate(unit = "Zipcode")
```

```{r}
bind_rows(cnts_state_avg_se, zcta_state_avg_se) %>%
  ggplot(aes(x = fct_reorder(state, avg), y = avg, ymax = avg + 1.96*se, ymin = avg - 1.96*se, fill = unit)) +
  geom_col(position = position_dodge()) +
  geom_errorbar(position = position_dodge()) +
  coord_flip() +
  labs(x = "",
       y = "Average civic opportunity index (1-5)",
       fill = "Aggregation unit")

ggsave(here("plots", "civic_opp_index_plots.png"))
```

```{r}

```

# Predictive validity 

- Poverty
- College education
- Race
- Gender
- Age
- Immigration
- Urbanicity

```{r}
corr_mod <- dd %>%
  # group_by(State) %>%
  do(tidy(standardize(lm_robust(formula = opc ~ race_per_white_nonhispanic + per_poverty + college_educ, data = .)))) # Here standardize refers to refitting the model with standardized data (https://cran.r-project.org/web/packages/effectsize/vignettes/effectsize.html) z-scoring

corr_mod_rc <- dd %>%
  # group_by(State) %>%
  do(tidy(standardize(lm_robust(formula = all_org ~ race_per_white_nonhispanic + per_poverty + college_educ, data = .))))
```

```{r}
corr_mod_bind <- bind_rows(
  mutate(corr_mod, DV = "Civic opportunity index"),
  mutate(corr_mod_rc, DV = "Organizational density")
)

pred_plot <- corr_mod_bind %>%
  mutate(term = recode(term,
    "race_per_white_nonhispanic" = "Non-hispanic white",
    "per_poverty" = "Federal poverty level",
    "college_educ" = "College educated"
  )) %>%
  filter(!str_detect(term, "(Intercept)")) %>%
  ggplot(aes(x = term)) +
  geom_text_repel(aes(y = estimate, label = round(estimate, 2), fill = DV, size = 1.5),
    position = position_dodge(width = 0.9),
    col = "black",
    show.legend = FALSE
  ) +
  geom_col(aes(y = estimate, fill = DV), position = "dodge", alpha = 0.4) +
  geom_errorbar(aes(ymax = estimate + std.error, ymin = estimate - std.error, fill = DV), position = "dodge", alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed", col = "red") +
  labs(
    x = "Predictors",
    y = "Estimated regression coefficient",
    fill = "Response variables"
  ) +
  scale_fill_viridis_d(begin = 0.1, end = 0.9) +
  # facet_wrap(~term, ncol = 2) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  coord_flip() +
  theme(legend.position = "bottom")

pred_plot

ggsave(here("outputs", "pred_plot.png"), height = 6, width = 6)
```